Pre-processing high frequency animal tracking data

### 2020-10-02 05_calibration_data.Rmd
some new figures

```{r}
# read and plot example data
data <- fread("data/atlas1060_allTrials_annotated.csv")

# plot data
fig_data_raw <- ggplot(data)+
  geom_path(aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(aes(x, y),
             col = "steelblue", alpha = 0.2, size = 2)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

Save a raw copy.

```{r}
data_raw <- copy(data)
```

@@

```{r}
# plot data
fig_data_bbox <-
  ggplot()+
  geom_path(data = data_raw,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_raw[!data, on = c("x", "y")],
             aes(x, y),
             col = "red", alpha = 1, size = 2)+
  geom_point(data = data,
             aes(x, y),
             col = "steelblue", alpha = 0.2, size = 2)+
  geom_vline(xintercept = 645000,
             col = "grey", lty = 2)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

```{r}
# save a copy
data_raw <- copy(data)

@@

```{r}
fig_low_nbs <- ggplot()+
  geom_path(data = data,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_raw[!data, on = c("x", "y")],
            aes(x, y),
            shape = 4,
            col = "red", alpha = 0.5, size = 0.3)+
  geom_point(data = data,
             aes(x, y, col = NBS),
             alpha = 1, size = 0.2)+
  scico::scale_colour_scico(palette = "davos",
                         labels = scales::comma, end = 0.9)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)+
  labs(colour = "NBS")

@@

```{r}
# plot filtered data
fig_speed_outliers <-
  ggplot()+
  geom_path(data = data,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data,
             aes(x, y, col = speed_in),
             size = 0.2)+
  scico::scale_colour_scico(palette = "davos", 
                         direction = 1, 
                         limits = c(NA, speed_angle_thresholds[1]),
                         na.value = "red",
                         labels = scales::comma,
                         end = 0.9,
                         trans = "sqrt")+
  ggthemes::theme_map()+
  theme(legend.position = c(0, 0.8))+
  coord_sf(crs = 32631)+
  labs(colour = "speed (m/s)")

@@

```{r}
# make a copy
data_raw <- copy(data)

@@

```{r}
# apply a 5 point median smooth, first make a copy
data_raw <- copy(data)

@@

```{r}
# make figure
fig_speed_outliers <-
  ggplot()+
  geom_path(data = data_raw,
            aes(x, y, col = "red"), 
            alpha = 0.5, size = 4)+
  geom_path(data = data,
             aes(x, y, col = "black"))+
  scico::scale_colour_scico_d(palette = "roma",
                         direction = -1, begin = 0.1,
                         labels = c("smoothed", "raw"))+
  ggthemes::theme_map()+
  theme(legend.position = c(0, 0.8))+
  coord_sf(crs = 32631)+
  labs(colour = "track")

# save figure
ggsave(fig_speed_outliers, 
       filename = "figures/fig_calib_median_smooth.png",
       width = 185 / 25, height = 185 / 25)
```

@@

```{r}
# make a data copy
data_raw <- copy(data)

@@

# get a simple inverse of speed as residence time classification
data_slow <- copy(data_raw)
data_slow <- data_slow[speed_in < 1, ]
```

@@

@@

```{r}
# patch with residence points and walking patches
fig_basic_residence <-
  ggplot()+
  geom_point(data = data,
             aes(x, y,
                 col = residence),
            size = ifelse(data$residence, 4, 0.2),
            shape = ifelse(data$residence, 2, 1),
            show.legend = FALSE)+
  scico::scale_color_scico_d(palette = "berlin",
                             direction = -1,
                             begin = 0.2, end = 0.8)+
  ggthemes::theme_map()

# figure of residence patches known
fig_patch_known <- 
  ggplot()+
  geom_sf(data = patch_sf_data[[1]],
          fill = "steelblue", col = NA)+
  geom_path(data = patch_summary_data[[1]],
            aes(x_median, y_median),
            col = "red")+
  ggthemes::theme_map()

@@

```{r}
# make figure comparing different methods
ggplot()+
  geom_point(data = patch_summary_compare,
             aes(nfixes_real, 
                 nfixes),
             show.legend = F,
            size = 4,
            shape = 1)+
  geom_smooth(data = patch_summary_compare,
             aes(nfixes_real, 
                 nfixes),
             show.legend = F)+
  scale_x_log10()+
  scale_y_log10()+
  coord_cartesian(xlim = c(10, NA))+
  ggthemes::theme_few()+
  labs(x = "# positions (real)",
       y = "# positions (inferred)")
```

>>>>

```{r}
# read and plot example data
data <- fread("data/atlas1060_allTrials_annotated.csv")
data_raw <- copy(data)
# plot data
fig_data_raw <- ggplot(data)+
  geom_path(aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(aes(x, y),
             col = "steelblue", alpha = 0.2, size = 2)+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

Save an unprocessed copy.

```{r}
data_unproc <- copy(data)
```

@@

```{r}
# plot data
fig_data_bbox <-
  ggplot()+
  geom_path(data = data_unproc,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_unproc,
             aes(x, y,
                 col = x > 645000),
             alpha = ifelse(data_unproc$x > 645000, 0.5, 1),
             size = ifelse(data_unproc$x > 645000, 0.3, 2),
             show.legend = F)+
  scale_color_brewer(palette = "Set1")+
  geom_vline(xintercept = 645000,
             col = "grey", lty = 2)+
  ggspatial::annotation_scale(location = "br")+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

@@

```{r}
# save a copy
data_unproc <- copy(data)

@@

```{r}
fig_low_nbs <-
  ggplot()+
  geom_path(data = data,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data_unproc,
            aes(x, y,
                col = as.factor(NBS > 3)),
            shape = 1,
            alpha = 0.5, size = 0.3, show.legend = F)+
  scale_color_brewer(palette = "Set1")+
  ggthemes::theme_map()+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(colour = "NBS")

@@

```{r}
# plot filtered data
fig_speed_outliers <-
  ggplot()+
  geom_path(data = data,
            aes(x, y),
             col = "grey", alpha = 0.5, size = 0.3)+
  geom_point(data = data,
             aes(x, y, 
                 col = speed_in < 15.7),
             size = 0.3, alpha = 0.5,
             show.legend = F)+
  scale_color_brewer(palette = "Set1")+
  ggthemes::theme_map()+
  ggspatial::annotation_scale(location = "tl")+
  coord_sf(crs = 32631)+
  labs(colour = "speed (m/s)")

@@

```{r}
# make a copy
data_unproc <- copy(data)

@@

```{r}
# apply a 5 point median smooth, first make a copy
data_unproc <- copy(data)

@@

```{r}
# make figure
fig_smooth <-
  ggplot()+
  geom_path(data = data,
             aes(x, y),
            col = RColorBrewer::brewer.pal(4, "Set1")[4])+
  annotate(geom = "rect",
           xmin = c(650785, 650045),
           ymin = c(5904450, 5902700),
           xmax = c(652500, 650682),
           ymax = c(5906133, 5903080),
           fill = "grey90",
           alpha = 0.5,
           col = NA)+
    annotate(geom = "text",
           x = c(650785, 650045) + 400,
           y = c(5904450, 5902700) + 300,
           col = "black",
           label = c("(a)", "(b)"))+
  theme_void()+
  coord_sf(crs = 32631)+
  labs(colour = "track")

# make zoomed in figures
fig_inset <- mapply(function(xl, yl) {
  ggplot()+
    geom_point(data = data_raw[!data_unproc, on = c("x", "y")],
               aes(x, y),
               col = "grey",
               shape = 4,
               size = 0.4)+
    geom_point(data = data_unproc,
               aes(x, y),
               col = RColorBrewer::brewer.pal(3, "Set1")[3],
               shape = 1, 
               alpha = 0.5)+
    geom_path(data = data,
              aes(x, y),
              col = RColorBrewer::brewer.pal(4, "Set1")[4])+
    ggthemes::theme_few()+
    theme(axis.title = element_blank(),
          axis.text = element_blank())+
    ggspatial::annotation_scale(location = "br")+
    coord_sf(crs = 32631,
             xlim = xl,
             ylim = yl, expand = F
             )+
    labs(colour = "track")
}, list(xl1 = c(650785, 652500),
        xl2 = c(650045, 650682)),
   list(yl1 = c(5904450, 5906133),
        yl2 = c(5902700, 5903080)),
SIMPLIFY = FALSE)

# add global figure to second inset
fig_inset[[2]] <- fig_inset[[2]] +
  annotation_custom(grob = ggplotGrob(fig_smooth),
                    xmin = 649990,
                    xmax = 650045 + (650682 - 650045)/2,
                    ymin = 5903080 - (5903080 - 5902700)/2.5,
                    ymax = 5903080
                    )

wrap_plots(fig_inset)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")
# save figure
ggsave(filename = "figures/fig_calib_median_smooth.png",
       width = 185 / 25, height = 90 / 25)
```

@@

```{r}
# make a data copy
data_unproc <- copy(data)

@@

# get a simple inverse of speed as residence time classification
data_slow <- copy(data_unproc)
data_slow <- data_slow[speed_in < 1, ]
```

@@

# assign a crs
patch_sf_data <- lapply(patch_sf_data, function(df) {
  sf::st_crs(df) <- 32631
  return(df)
})

@@

```{r}
# patch with residence points and walking patches
fig_basic_residence <-
  ggplot()+
  geom_point(data = data,
             aes(x, y,
                 col = residence),
            size = ifelse(data$residence, 4, 0.2),
            shape = ifelse(data$residence, 2, 1),
            show.legend = FALSE)+
  scale_color_brewer(palette = "Set1")+
  ggthemes::theme_map()+
  coord_sf(crs = 32631)

# figure of residence patches known
# fig_patch_known <- 
  ggplot()+
  geom_sf(data = patch_sf_data[[1]],
          fill = "steelblue", col = NA)+
  geom_path(data = patch_summary_data[[1]],
            aes(x_median, y_median),
            col = "red")+
  ggthemes::theme_map()

pla = mapply(function(pol) {

  bbox = st_bbox(pol)

  ggplot()+
    geom_path(data = data_unproc,
               aes(x, y),
               col = RColorBrewer::brewer.pal(4, "Set1")[4])+
    geom_sf(data = pol,
            alpha = 0.2)+
    ggthemes::theme_map()+
    coord_sf(
      expand = F,
      xlim = c(bbox[["xmin"]], bbox[["xmax"]]),
      ylim = c(bbox[["ymin"]], bbox[["ymax"]]))
}, patch_sf_data[[1]]$polygons, SIMPLIFY = F)

fig_patches_small <- wrap_plots(pla)+
  plot_annotation(tag_levels = "a",
                  tag_prefix = "(",
                  tag_suffix = ")")

@@

```{r}
# make figure comparing different methods
ggplot()+
  geom_point(data = patch_summary_compare,
             aes(duration_real, 
                 duration),
             show.legend = F,
            size = 4,
            shape = 1)+
  geom_smooth(data = patch_summary_compare,
             aes(duration_real, 
                 duration),
             se = F,
             method = "glm",
             show.legend = F)+
  # scale_x_log10()+
  # scale_y_log10()+
  # coord_cartesian(xlim = c(10, NA))+
  ggthemes::theme_few()+
  labs(x = "# positions (real)",
       y = "# positions (inferred)")
```
